Restrictive Clearance Comparison

### 2024-04-18 data-raw/RestrictiveClearance_comparison.Rmd
An early beginning to a re-parametrized restrictive vs non-restrictive one-compartment model

@@

```{r running-invivoPKfit}
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))

@@

new_pk <- do_fit(new_pk) # For troubleshooting

new_pk <- do_fit(new_pk, n_cores = 10) 
# Multidplyr error because of forgotten browser() call within log_likelihood
new_pk$fit %>% unnest(fit) %>% View()

@@

```{r}

new_wm <- get_winning_model(new_pk)
new_AIC <- AIC(new_pk)
new_rsq <- rsq(new_pk)

@@

residuals(new_pk) %>% View()

pl <- plot(new_pk, n_interp = 12, use_scale_conc = TRUE, best_fit = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "BestFits_Final_2024_0405.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

```{r cltot-comparison}
new_cltot2 <- new_tkstats %>% 
  # filter(str_detect(model, "rest")) %>%
  select(Chemical, Species, Media, Route, 
         model, method,
         CLtot.tkstats, CLtot.nca) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(clearance_type = ifelse(str_detect(model, "nonrest"),
                                "NonRestrictive",
                                ifelse(str_detect(model, "rest"),
                                       "Restrictive",
                                       "optimized kelim"))) %>%
  rowwise() %>%
  mutate(CLtot.httk_rest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = TRUE),
         CLtot.httk_nonrest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = FALSE))

new_cltot2 %>%
  ggplot(aes(x = Chemical)) +
  geom_point(aes(y = CLtot.httk_rest),shape = 19, size = 2, color = "green4") +
  geom_point(aes(y = CLtot.httk_nonrest),shape = 19, size = 2, color = "orange2") +
  geom_point(aes(y = CLtot.tkstats), shape = 1, size = 3, color = "black", stroke = 1) +
  scale_y_log10(labels = scales::label_math(format = log10)) +
  geom_hline(yintercept = c(0.01, 0.1, 1, 10)) +
  coord_flip() +
  facet_grid(rows = vars(clearance_type),
             scales = "free", space = "free") +
  labs(title = "Total Clearance for constant kelim \"best fits\" (empty)\nand httk calculations (filled)") +
  theme_bw() +
  theme(axis.line = element_line(linewidth = 1),
        axis.text = element_text(size = 12),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_backcompare_2024_0405.png"),
       height = 4,
       width = 6)

@@

df_valcomp %>% 
  column_to_rownames(var = "Chemical") %>%
  t() %>%
  as.matrix() %>%
  cor() %>%
  round(digits = 3) -> val_cormat

hclust(d = as.dist(1-abs(val_cormat)), method = "complete") %>% plot()

hclust(d = as.dist(1-abs(val_cormat)), method = "complete") -> val_hclust

@@

#reorder the Chemical names prior to plotting
df_valcomp %>%
  mutate(Chemical = factor(x = Chemical,
                           levels = val_hclust$labels[val_hclust$order])) %>%
  pivot_longer(cols = A_rest:C_nonrest,
               names_to = "comparison", values_to = "logres") %>%
  mutate(comparison = case_when(
    comparison == "A_nonrest" ~ "free kelim vs httk nonrestrictive",
    comparison == "A_rest" ~ "free kelim vs httk restrictive",
    comparison == "B_nonrest" ~ "restrictive kelim vs httk nonrestrictive",
    comparison == "B_rest" ~ "restrictive kelim vs httk restrictive",
    comparison == "C_nonrest" ~ "nonrestrictive kelim vs httk nonrestrictive",
    comparison == "C_rest" ~ "nonrestrictive kelim vs httk restrictive",
  ),
  comparison = factor(comparison,
  levels = c("free kelim vs httk nonrestrictive",
             "free kelim vs httk restrictive",
             "restrictive kelim vs httk nonrestrictive",
             "restrictive kelim vs httk restrictive",
             "nonrestrictive kelim vs httk nonrestrictive",
             "nonrestrictive kelim vs httk restrictive"))) %>%
  ggplot(aes(x = comparison, y = Chemical, fill = logres)) +
  geom_tile() +
  scale_fill_gradient2(mid = "grey5", high = "magenta", low = "green",
                       name = "Total Clearance\nlog-scale difference") +
  coord_fixed(ratio = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        axis.title.x = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_compare_heatmap_2024_0405.png"),
       height = 14,
       width = 8)

@@

@@

>>>>

common_chems_dawson <- cvt %>%
  subset(analyte_dtxsid %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk")) %>%
  pull(analyte_dtxsid) %>%
  unique()

cvt %>%
  subset(!(analyte_dtxsid %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk"))) %>%
  dplyr::distinct(analyte_dtxsid, analyte_name_original)

# Test which analyte_dtxsid don't have a clearance or vdist

data.frame(Chemical = unique(cvt$analyte_dtxsid)) %>%
  rowwise() %>%
  mutate(needs_RestCL = ifelse(
    (class(try(calc_total_clearance(dtxsid = Chemical,
                     restrictive.clearance = TRUE,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE),
    needs_nonRestCL = ifelse(
      (class(try(calc_total_clearance(dtxsid = Chemical,
                     restrictive.clearance = FALSE,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE),
    needs_Vdist = ifelse(
      (class(try(calc_vdist(dtxsid = Chemical,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE)) %>%
  ungroup() -> has_all_needed_df

has_all_needed_df %>%
  filter(!(needs_RestCL | needs_nonRestCL | needs_Vdist)) %>%
  pull(Chemical) %>% unique() -> common_chems_dawson

@@

```{r running-invivoPKfit}
# This is now updated to use dawson chem table
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems_dawson),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))

@@

new_pk <- do_fit(new_pk) # For troubleshooting only

new_pk <- do_fit(new_pk, n_cores = 12)

# Multidplyr error because of forgotten browser() call within log_likelihood
new_pk$fit %>% unnest(fit) %>% View()

saveRDS(new_pk, "data-raw/someCVT_wHTTK_JOINT.rds")

@@

```{r new_analysis}
# Loading the new data
new_pk <- readRDS("data-raw/someCVT_wHTTK_JOINT.rds")

new_wm <- get_winning_model(new_pk)
new_AIC <- AIC(new_pk)
new_rsq <- rsq(new_pk)
new_rmse <- rmse(new_pk, use_scale_conc = FALSE)

@@

new_resid <- residuals(new_pk)

my_data <- get_data(new_pk) %>%
  distinct(Chemical, Species, Reference, Route, Media, Dose, Time, Conc) %>%
  right_join(new_resid)

pl <- plot(new_pk, n_interp = 12, use_scale_conc = FALSE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "AllFits_noScaleConc_newPK_2024_0412.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

```{r cltot-comparison}
new_cltot2 <- new_tkstats %>% 
  filter(str_detect(model, "rest")) %>%
  select(Chemical, Species, Media, Route, 
         model, method,
         CLtot.tkstats, CLtot.nca) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(clearance_type = ifelse(str_detect(model, "nonrest"),
                                "NonRestrictive",
                                ifelse(str_detect(model, "rest"),
                                       "Restrictive",
                                       "optimized kelim"))) %>%
  rowwise() %>%
  mutate(CLtot.httk_rest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = TRUE),
         CLtot.httk_nonrest = calc_total_clearance(dtxsid = Chemical,
                                           restrictive.clearance = FALSE))

new_cltot2 %>%
  ggplot(aes(x = Chemical)) +
  geom_point(aes(y = CLtot.httk_rest),shape = 19, size = 2, color = "green4") +
  geom_point(aes(y = CLtot.httk_nonrest),shape = 19, size = 2, color = "orange2") +
  geom_point(aes(y = CLtot.tkstats, shape = Species), size = 3, color = "black", stroke = 1) +
  scale_y_log10(labels = scales::label_math(format = log10)) +
  geom_hline(yintercept = c(0.01, 0.1, 1, 10)) +
  scale_shape_manual(values = c(0,1,2,5,6)) +
  coord_flip() +
  facet_grid(rows = vars(clearance_type),
             scales = "free", space = "free") +
  labs(title = "CLtot for constant kelim \"best fits\" (empty)\nand httk calculations (filled)") +
  theme_bw() +
  theme(axis.line = element_line(linewidth = 1),
        axis.text = element_text(size = 12),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0),
        strip.background = element_blank(),
        legend.position = "none")

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_backcompare_wDawson_2024_0416.png"),
       height = 8,
       width = 9)

@@

df_valcomp %>% 
  column_to_rownames(var = "Chemical") %>%
  t() %>%
  as.matrix() %>%
  cor() %>%
  signif() -> val_cormat

hclust(d = as.dist((1+val_cormat)/2), method = "average") %>% plot()

hclust(d = as.dist((1+val_cormat)/2), method = "average") -> val_hclust

@@

#reorder the Chemical names prior to plotting
df_valcomp %>%
  mutate(Chemical = factor(x = Chemical,
                           levels = val_hclust$labels[val_hclust$order])) %>%
  pivot_longer(cols = A_rest:C_nonrest,
               names_to = "comparison", values_to = "logres") %>%
  mutate(comparison = case_when(
    comparison == "A_nonrest" ~ "free kelim vs httk nonrestrictive",
    comparison == "A_rest" ~ "free kelim vs httk restrictive",
    comparison == "B_nonrest" ~ "restrictive kelim vs httk nonrestrictive",
    comparison == "B_rest" ~ "restrictive kelim vs httk restrictive",
    comparison == "C_nonrest" ~ "nonrestrictive kelim vs httk nonrestrictive",
    comparison == "C_rest" ~ "nonrestrictive kelim vs httk restrictive",
  ),
  comparison = factor(comparison,
  levels = c("free kelim vs httk restrictive",
             "free kelim vs httk nonrestrictive",
             "restrictive kelim vs httk restrictive",
             "nonrestrictive kelim vs httk nonrestrictive",
             "nonrestrictive kelim vs httk restrictive",
             "restrictive kelim vs httk nonrestrictive"))) %>%
  ggplot(aes(x = comparison, y = Chemical, fill = logres)) +
  geom_tile() +
  scale_fill_gradient2(mid = "grey5", high = "orange", low = "green3",
                       name = "Total Clearance\nlog10-scale difference",
                       limits = c(-4, 4),
                       oob = scales::oob_squish) +
  coord_fixed(ratio = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        axis.title.x = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_compare_heatmap_2024_0409.png"),
       height = 14,
       width = 8)

@@

# Winning models - An aside

I am a little worried that I am not selecting the best fit model for each
chemical all the time. Ultimately, the consequences for this being having parameters
for chemicals with rather poor fits in general, when there is a better fitting
model available. Here I want to do a quick analysis of the chemical
*DTXSID3022405* in mouse. The one compartment fits best in the original data
(by visual inspection) but the 2-compartment wins out in minimizing AIC.

For each model, I can find how many distinct experimental timepoints have the minimum
RMSE then investigate whether the model with the most _minimum RMSE_ timepoint
corresponds to the winning model by AIC.

*Update*: About half do not correspond. However, the lack of correspondance doesn't
mean there is bad fits declared winning model. To address this I need to understand 
_how_ different the AIC-based winning model and the RMSE-based winning models
are and whether the data leaves something to be desired.
This may be as simple as a Pearson correlation between residual values.
The output of this analysis is how related both

Another thing I am noticing in oral data: the constant _kelim_ fits sometimes are very similar
to their respective non-constant _kelim_ fits and they actually intersect/cross over.
I wonder if there is something special about the timepoints at which they intersect.
For example, a restrictive fit will reach a slightly higher Cmax, but then
dive below the fully optimized fit during elimination phase before crossing over
again near the second inflection point.

```{r oxybenzene-test}
old_AIC <- AIC(my_pk) %>%
  filter(Chemical %in% "DTXSID3022405",
         Species %in% "mouse")

new_AIC <- new_AIC %>%
  filter(Chemical %in% "DTXSID3022405",
         Species %in% "mouse")

plot(new_pk, model = c("model_2comp",
                       "model_1comp", "model_1comp_nonrest"),
     use_scale_conc = FALSE) %>%
  filter(Chemical %in% "DTXSID3022405",
         Species %in% "mouse") %>%
  pull(final_plot) %>% .[[1]]

old_rsquared <- rmse(my_pk) %>%
  filter(Chemical %in% "DTXSID3022405",
         Species %in% "mouse")

# Only test residual correlations with those not the same

winmodel_conflict <- rmse(new_pk, use_scale_conc = FALSE) %>%
  group_by(Chemical, Species, Route, Media, Dose, Time) %>%
  filter(RMSE == min(RMSE)) %>%
  group_by(Chemical, Species) %>%
  count(model) %>%
  filter(n == max(n)) %>%
  inner_join(new_wm %>% rename("winmodel" = "model")) %>%
  filter(!(model == winmodel))

# need to map2 then have `residuals.pk(model = .x)` and .y
# then correlate these values for each data group
winmodel_conflict %>% names()

mod_red <- winmodel_conflict %>%
  dplyr::select(Chemical, Species, model) %>%
  left_join(new_resid,
            by = join_by(Chemical, Species, model)) %>%
  rename("mod_residuals" = "Residuals") %>%
  select(Chemical, Species, Route, Media, Time, Detect, exclude,
         model, mod_residuals)

wm_red <- winmodel_conflict %>%
  dplyr::select(Chemical, Species, winmodel) %>%
  left_join(new_resid,
            by = join_by(Chemical, Species, "winmodel" == "model")) %>%
  rename("wm_residuals" = "Residuals") %>%
  select(Chemical, Species, Route, Media, Time, Detect, exclude,
         winmodel, wm_residuals)
sum_red <- new_resid %>%
  group_by(Chemical, Species, model) %>%
  summarize(MSE = sum(Residuals^2)/n())

cor_mod <- mod_red %>% inner_join(wm_red) %>%
  select(Chemical, Species, model, mod_residuals, winmodel, wm_residuals) %>%
  group_by(Chemical, Species, model, winmodel) %>%
  summarize(cor_mods = cor(x = mod_residuals, y = wm_residuals)) %>%
  left_join(sum_red, by = join_by(Chemical, Species, model)) %>%
  rename("mod_MSE" = "MSE") %>%
  left_join(sum_red, by = join_by(Chemical, Species, "winmodel" == "model")) %>%
  rename("wm_MSE" = "MSE")

my_pl_1comp <- plot(new_pk, n_interp = 11, use_scale_conc = FALSE,
                    model = c("model_1comp",
                              "model_1comp_rest",
                              "model_1comp_nonrest"))

my_pl_2comp <- plot(new_pk, n_interp = 11, use_scale_conc = FALSE,
                    model = c("model_2comp",
                              "model_2comp_rest",
                              "model_2comp_nonrest"))

# General trends about residual correlation and MSE
cor_mod %>%
  mutate(cor_bin = ifelse(abs(cor_mods) >= 0.60,
                          "Taking win_model",
                          "Taking low RMSE model")) %>% na.omit() %>%
  ggplot() +
  geom_point(aes(x = mod_MSE, y = wm_MSE)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  facet_grid(rows = vars(cor_bin))

cor_mod %>%
  filter(!(model %in% "model_flat")) %>%
  mutate(cor_bin = ifelse(abs(cor_mods) >= 0.60,
                          "Taking win_model",
                          "Taking low RMSE model")) %>% na.omit() %>%
  ggplot() +
  geom_histogram(aes(x = cor_mods, fill = model),
                 binwidth = 0.05) +
  facet_grid(rows = vars(winmodel)) +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0)) +
  labs(x = "Correlation of Residuals",
       fill = "model by grouped RMSE")

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "winModel_correlations_changes",
                         "_2024_Apr_16.png"),
       height = 6,
       width = 8)

# Try oxybenzone fitting without dose-norm
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% "DTXSID3022405"),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
new_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference)) +
  stat_model(model = c("model_flat", "model_1comp", "model_2comp",
                       "model_1comp_rest", "model_1comp_nonrest",
                       "model_2comp_rest", "model_2comp_nonrest"))

new_pk <- do_fit(new_pk, n_cores = 3) 
get_winning_model(new_pk)
plot(new_pk, best_fit = TRUE) %>% pull(final_plot)
```

# Categorizing chemicals

I first need to find the winning model for constant kelims and optimized kelims

```{r categorization}

grouped_wm <- new_AIC %>%
  mutate(model_type = case_when(
    str_detect(model, "_nonrest$") ~ "Nonrestrictive",
    str_detect(model, "_rest$") ~ "Restrictive",
    .default = "Free_kelim"
  )) %>%
  group_by(Chemical, Species, model_type) %>%
  filter(AIC == min(AIC)) %>% 
  ungroup() %>%
  arrange(Chemical, Species, model_type) %>%
  select(Chemical, Species, model, method, model_type)

grouped_tkstats <- get_tkstats(new_pk) %>%
  inner_join(grouped_wm)

grouped_tkstats %>%
  select(Chemical, Species, Reference,
         Route, Media, model_type,
         CLtot) %>%
  pivot_wider(names_from = model_type, values_from = CLtot) %>%
  mutate(low_invivo_CL = ifelse(
    Free_kelim < Nonrestrictive/2 | Free_kelim < Restrictive/2,
    TRUE, FALSE),
    high_extra_hepCL = ifelse(
      Free_kelim > 1.5*Nonrestrictive | Free_kelim > 1.5*Restrictive,
      TRUE, FALSE),
    rest_metabol = ifelse(
      Free_kelim > Restrictive/2 & Free_kelim < 1.5*Nonrestrictive,
      TRUE, FALSE),
    combo_cat = paste0(low_invivo_CL + 0,
                       high_extra_hepCL + 0,
                       rest_metabol + 0)) %>%
  count(combo_cat)
  arrange(Chemical, combo_cat) %>%
  na.omit() %>%
  View()

@@

```{r}
my_tkstats <- eval_tkstats(my_pk, finite_only = FALSE)

my_tkstats %>% filter(Species %in% "rat",
                      Route %in% "oral") %>%
  distinct(Chemical, Species, Route, Media, model, Fgutabs.tkstats) %>%
  filter(Fgutabs.tkstats > 0) %>%
  left_join(distinct(cas_convert)) %>% 
  arrange(Chemical) %>% clipr::write_clip()

pesticides <- read_csv("C:/Users/GPADILLA/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/Chemical List EPAPCS-2024-04-17.csv") %>%
  pull(DTXSID) %>% str_extract("DTXSID\\d+")

unique(my_tkstats$Chemical)[unique(my_tkstats$Chemical) %in% pesticides]
